---
title: "Lab 9"
author: "Jun Noh"
execute: 
  error: true
format: 
  html:
    self-contained: true
    toc: true
    toc-location: left
    code-fold: true
editor: visual
---

## Loading in libraries/dataset

```{r}
#| message: false
library(tidyverse)
library(here)
library(broom)
```

```{r}
#| message: false
babyNames <- read_csv(here::here("supporting_artifacts", "Learning Targets", 
                                 "lab9", "StateNames_A.csv")) |> 
  rename(Sex = "Gender")
```

## **1. Summarizing & Visualizing the Number of Allisons**

```{r}
#| message: false
babyNames |> 
  mutate(Count = replace_na(Count, 0)) |> 
  group_by(Sex, State) |> 
  filter(Name == "Allison", Sex == "F") |> 
  summarize(total = sum(Count))
```

## 1a.

```{r}
allisons <- babyNames |> 
  group_by(Year) |> 
  filter(Name == "Allison", Sex == "F") |> 
  summarize(total = sum(Count)) 
```

## 1b.

```{r}
babyNames |> 
  group_by(Year) |> 
  filter(Name == "Allison", Sex == "F") |> 
  summarize(total = sum(Count)) |> 
  ggplot(mapping = aes(x = Year, y = total)) +
    geom_line() +
    labs(x = "Year", title = "Number of babies named Allison by year (Females only)") +
    theme(axis.title.y = element_blank()) +
    scale_x_continuous(limits = c(1997,2015))
```

# 2. Modeling the number of Allisons

```{r}
babyNames_lm <- allisons |> 
  lm(total ~ Year, data = _)
```

```{r}
#| message: false
allisons |> 
  ggplot(mapping = aes(x = Year, y = total)) + 
    geom_line() +
    stat_smooth(method = "lm") +
    labs(x = "Year", title = "Linear Model of Number of Allisons born per Year") +
    theme(axis.title.y = element_blank()) 
```

## 3. Regression Model

```{r}
tidy(babyNames_lm)
```

The predicted number of names of Allsions by year can be represented with the expression below.

Predicted number of Allisons = 209,689.7609 - 101.5191(Year)

## 4. Plotting Resiudals

```{r}
babyNames_lm |> 
  augment() |> 
  ggplot(mapping = aes(x = Year, y = .resid)) + 
    geom_line() + 
    labs(x = "Year", title = "Residuals by Year for the name Allison") +
    theme(axis.title.y = element_blank()) 
```

The residuals increase slightly from 1997 to 1998 then steadily decrease after until 2008 where it begins to pick up again. There is a slight decrease again in 2011-2012 followed by a slight increase. Based on the graphs slight upward trend I believe the name Allison is somewhat popular now, but nowhere as popular as it once was in 1998.

## 5. Spelling by State

```{r}
A_names <- babyNames |> 
  filter(Name %in% c("Alan", "Allen", "Allan"), Sex == "M")
```

```{r}
#| message: false
A_names |> 
  group_by(Year, Name) |> 
  summarize(total = sum(Count)) |> 
  ggplot(mapping = aes(x = Year, y = total, 
                       color = fct_reorder2(Name, Year, total))) +
    geom_line() + 
    labs(x = "Year", title = "Number of babies by year and spelling of Alan", 
         color = "Spelling of Alan") +
    theme(axis.title.y = element_blank()) 
```

```{r}
#| message: false
A_names |> 
  filter(State %in% c("CA", "PA"), Year == "2000") |> 
  pivot_wider(names_from = Name, values_from = Count) |> 
  mutate(across(Alan:Allan, .fns = replace_na, 0)) |> 
  rowwise() |> 
  mutate(total = sum(Alan, Allen, Allan))
```

```{r}
#| message: false
A_names |> 
  filter(State %in% c("CA", "PA"), Year == "2000") |> 
  pivot_wider(names_from = Name, values_from = Count) |> 
  mutate(across(Alan:Allan, .fns = replace_na, 0)) |> 
  rowwise() |> 
  mutate(total = sum(Alan, Allen, Allan), Alan = Alan/total * 100, 
         Allen = Allen/total * 100, Allan = Allan/total * 100, 
         total = sum(Alan, Allen, Allan)) 
```
